<<<<<<< HEAD
---
title: "st542_project10"
author: "Jingjing Li"
date: "2023-06-15"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(haven)
library(psych)
```
### Exploring the merged data

```{r}
survey_merged <- read_dta("../data/merged_survey_panel96_19_noID.dta") 
dim(survey_merged)
survey_merged_clean <- survey_merged %>% 
  filter(!is.na(lotid_qualtrics))
# Removing duplicates for modelling - forgot to ask Mariana which should be kept, so just keeping first for now
# Also open to just dropping all duplicated lotids - there's only 42
survey_final <- survey_merged_clean %>% distinct(lotid_qualtrics, .keep_all = TRUE)
```

Now to check for nulls in columns:
```{r}
missing_vals <- survey_final %>% summarise_all(~sum(is.na(.)))
t(missing_vals)
```
There seem to be a lot of columns missing 1026 values, which puts the numbers back to what we saw in the original data. Clearly there is a lot of missing data across all areas, so I don't think we should drop all rows missing data. 

## Feature Engineering

### Creating One Drainage Area Column 
I will create one drainage area column, and then will drop all rows that don't have drainage area. Not sure if this is the best approach but this data is so messy!

```{r}
# There are two lots that have areas in each region - will drop them
duplicate_regions <- survey_final %>% filter(!is.na(OPO_area_km) & !is.na(RO_area_km)) %>% select(lotid_qualtrics)
survey_final <- survey_final %>% filter(!(lotid_qualtrics %in% duplicate_regions$lotid_qualtrics))
survey_final <- survey_final %>% mutate(drainage_area_km = ifelse(!is.na(Drainage_AreaKM_Ari), Drainage_AreaKM_Ari, ifelse(!is.na(OPO_area_km), OPO_area_km, ifelse(!is.na(RO_area_km), RO_area_km, NA))),
                                        region = ifelse(studycode == 1, "Ariquemes",
                                                        ifelse(studycode == 2, "Ouro Preto do Oeste", ifelse(studycode == 3, "Rolim de Moura", NA)))) %>%
  select(-c(Drainage_AreaKM_Ari, RO_area_km, OPO_area_km))
```


### Creating Individual Adaptation Method Booleans
First, I need to create the individual adaptation method booleans i.e. `cattle_management` if the farmer employed any cattle management strategy. Then I can combine these to make a general adaptation method boolean.

```{r}
survey_final <- survey_final %>% mutate(cattle_management = case_when(feed_cattle == 1  | soil_analysis == 1 | cattleinputs_breed == 1 | semiconfine == 1 | sellcattle_droughtexp == 1 ~ 1, .default = 0),
                                        pasture_management = case_when(irrigation_pas == 1 | pasture_productivity == 1 | fert_pasture ==1 | pest_pasture == 1 | fallow == 1 ~ 1, .default = 0),
                                        forest_conservation = case_when(forest_rec == 1 | keep_veg == 1 ~ 1, .default = 0),
                                        water_management = case_when(waterstructure == 1 | well_have == 1 | dam_have == 1 | reservoir_have == 1 ~ 1, .default = 0)) # Chose to use case when to account for NA - need to ask Mariana how these should be treated

# Looking at distributions
g <- ggplot(data = survey_final, aes(x = cattle_management))
g + geom_bar() + labs(title = "Bar Plot of Cattle Management")

g <- ggplot(data = survey_final, aes(x = pasture_management))
g + geom_bar() + labs(title = "Bar Plot of Pasture Management")

g <- ggplot(data = survey_final, aes(x = forest_conservation))
g + geom_bar() + labs(title = "Bar Plot of Forest Conservation")

g <- ggplot(data = survey_final, aes(x = water_management))
g + geom_bar() + labs(title = "Bar Plot of Water Management")

```

### Creating General Adaptation Method Boolean
Now that the indivdual adaptation measures have been created, I will create a general adaptation method boolean

```{r}
survey_final <- survey_final %>% mutate(any_adaptation = ifelse(cattle_management == 1 | pasture_management == 1 | forest_conservation == 1 | water_management == 1,1,0))

# Looking at the distribution
g <- ggplot(data = survey_final, aes(x = any_adaptation))
g + geom_bar() + labs(title = "Bar Plot of Any Adaptation")
```
Not a fully even data set, but better balanced than I predicted it would be. I don't
think rebalancing is needed for the final analysis.

Now let's see if we remove farmers who don't have GIS information
```{r}
survey_final_gis <- survey_final %>% filter(!is.na(drainage_area_km))

# Looking at the distribution
g <- ggplot(data = survey_final_gis, aes(x = any_adaptation))
g + geom_bar() + labs(title = "Bar Plot of Any Adaptation")
```
Now it is very imbalanced. Not as horrible as it could be - but could affect the outputs of the model. Is SPI still missing for these?

```{r}
missing_vals_gis <- survey_final_gis %>% summarise_all(~sum(is.na(.)))
t(missing_vals_gis)
```
Now only for 140 of them, who are also missing region. Let's drop those and see:
```{r}
survey_final_gis_clean <- survey_final_gis %>% filter(!is.na(SPImin_year))

# Looking at the distribution
g <- ggplot(data = survey_final_gis_clean, aes(x = any_adaptation))
g + geom_bar() + labs(title = "Bar Plot of Any Adaptation")

# Looking at the distribution for the individual variables
g <- ggplot(data = survey_final_gis_clean, aes(x = cattle_management))
g + geom_bar() + labs(title = "Bar Plot of Cattle Management")

g <- ggplot(data = survey_final_gis_clean, aes(x = pasture_management))
g + geom_bar() + labs(title = "Bar Plot of Pasture Management")

g <- ggplot(data = survey_final_gis_clean, aes(x = forest_conservation))
g + geom_bar() + labs(title = "Bar Plot of Forest Conservation")

g <- ggplot(data = survey_final_gis_clean, aes(x = water_management))
g + geom_bar() + labs(title = "Bar Plot of Water Management")
```
And everyone did some sort of adaptation method. Great - this means if we drop all missing water data, we can't look at a relationship between general adaptation and results. If we include rows that are missing SPI and just use drainage area, there is still a model that can be run.

## Exploratory Data Analysis

Start EDA with some individual variable histograms.

```{r}
g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km))
g + geom_histogram() + labs(title = "Histogram of Drainage Area")

g <- ggplot(data = survey_final_gis, aes(x = rainfall,y=any_adaptation))
g + geom_histogram() + labs(title = "Boxplot of Rainfall")

g <- ggplot(data = survey_final_gis, aes(x = SPI_year))
g + geom_histogram() + labs(title = "Histogram of Average Yearly SPI")

g <- ggplot(data = survey_final_gis, aes(x = vechval))
g + geom_histogram() + labs(title = "Histogram of Vehicle Value")

g <- ggplot(data = survey_final_gis, aes(x = family))
g + geom_histogram() + labs(title = "Histogram of Family Size")

g <- ggplot(data = survey_final_gis, aes(x = lotprice))
g + geom_histogram() + labs(title = "Histogram of Lot Price")

```
Next, we can look at some box plots
```{r}
g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km))
g + geom_boxplot() + labs(title = "Boxplot of Drainage Area")

g <- ggplot(data = survey_final_gis, aes(x = rainfall,y=any_adaptation))
g + geom_boxplot() + labs(title = "Boxplot of Rainfall")

g <- ggplot(data = survey_final_gis, aes(x = SPI_year))
g + geom_boxplot() + labs(title = "Boxplot of Average Yearly SPI")

g <- ggplot(data = survey_final_gis, aes(x = SPI_dry))
g + geom_boxplot() + labs(title = "Boxplot of Average in Peak Dry Season")

g <- ggplot(data = survey_final_gis, aes(x = SPI_wet))
g + geom_boxplot() + labs(title = "Boxplot of Average in Peak Wet Season")

g <- ggplot(data = survey_final_gis, aes(x = vechval))
g + geom_boxplot() + labs(title = "Boxplot of Vehicle Value")

g <- ggplot(data = survey_final_gis, aes(x = family))
g + geom_boxplot() + labs(title = "Boxplot of Family Size")

g <- ggplot(data = survey_final_gis, aes(x = lotprice))
g + geom_boxplot() + labs(title = "Boxplot of Lot Price")


```


Let's look at boxplots of drainage area and vehicle value (wealth) grouped by cattle.

```{r}
g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km, y = as.factor(havecattle)))
g + geom_boxplot() + labs(title = "Boxplot of Drainage Area grouped by Cattle") +
  ylab("Cattle")

g <- ggplot(data = survey_final_gis, aes(x = vechval, y = as.factor(havecattle)))
g + geom_boxplot() + labs(title = "Boxplot of Vehicle Value grouped by Cattle") +
  ylab("Cattle")

```

Let's look at a few box plots split by region

```{r}

g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Drainage Area grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = SPI_dry, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Dry Season SPI grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = SPI_wet, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Wet Season SPI grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = rainfall_dry, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Dry Season Rainfall grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = rainfall_wet, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Wet Season Rainfall grouped by Region") +
  ylab("Region")

```

Next lets look at pair plots

```{r}
pairs_subset <- survey_final_gis %>% select(c(drainage_area_km, SPImax_year, rainfall, studycode, havecattle, vechval, lotprice, incbeef))
pairs(pairs_subset)

```

Let's look at how many households fall into each region

```{r}
table(survey_final_gis$studycode)
```

Let's also look at some basic summary statistics

```{r}
describe(survey_final_gis[c('drainage_area_km', 'rainfall', 'rainfall_wet', 'rainfall_dry', 'SPI_wet', 'SPI_dry', 'incbeef', 'vechval', 'lotprice', 'family')], fast=TRUE)

```


I'm curious about how many farms made multiple adaptations

```{r}
table(survey_final_gis$cattle_management, survey_final_gis$pasture_management)

table(survey_final_gis$cattle_management, survey_final_gis$forest_conservation)

table(survey_final_gis$cattle_management, survey_final_gis$water_management)

table(survey_final_gis$pasture_management, survey_final_gis$forest_conservation)

table(survey_final_gis$pasture_management, survey_final_gis$water_management)

table(survey_final_gis$forest_conservation, survey_final_gis$water_management)
```

It looks like the majority of farmers who adapted did more than one thing.

# Statisitcal analysis

## t test for milkincome
We would like to see if there is any difference of milkincome between dry and wet season. The null hypothesis is mean value of milkincome is the same for dry and wet season. H0: $\mu_{mlkinc,dry}=\mu_{mlkinc,wet)}$ vs Ha: $\mu_{mlkinc,dry} \ne \mu_{mlkinc,wet)}$  

```{r}
t.test(survey_final_gis_clean$milkincwet,survey_final_gis$milkincdry,na.rm=TRUE,var.equal = FALSE)
```
As p is much less than 0.05, we are confident the mean value of milkincome of wetseason is significant higher than dry season.  

# t-test for SPI 
We would like to see if there is any difference in SPI between dry and wet season. The null hypothesis is mean value of SPI is the same for dry and wet season. H0: $\mu_{SPI,dry}=\mu_{SPI,wet)}$ vs Ha: $\mu_{SPI,dry} \ne \mu_{SPI,wet)}$ 
```{r}
t.test(survey_final_gis_clean$SPI_wet6,survey_final_gis$SPI_dry6,na.rm=TRUE,var.equal = FALSE)
t.test(survey_final_gis_clean$SPI_wet,survey_final_gis$SPI_dry,na.rm=TRUE,var.equal = FALSE)
```
Average SPI and average monthly SPI in the peak of wet  season is significantly higher than those of dry season.   

## Anova analysis
We hypothesize drainage area of three regions may be different. We propose null hypothesis is drainage area means of three regions are the same. The alternativae hypothesis is drainage area means of three regions are different.  

### effect of region on drainage_area_km
```{r}
survey_final_gis$studycode<- factor(survey_final_gis$studycode)
aov_drainage <- aov(drainage_area_km ~ studycode, data=survey_final_gis)

summary(aov_drainage)
TukeyHSD(aov_drainage, conf.level=.95)
```
As p-value of F test is close 0.05,  the drainage area means of three regions are somehow different. We further compared drainage means of three regions and found region3 (Rolim) has the highest drainage area, OPO has the smallest.  

### effect of region on income from beef
We then wants to see if there is difference in income from beef at different regions.
```{r}
aov_incbeef <- aov(incbeef ~ studycode, data=survey_final_gis)
summary(aov_incbeef)
TukeyHSD(aov_incbeef, conf.level=.95)
```
As p value is larger than 0.05, we can not reject our hypothesis that the means of income from beef at different regions are the same.  


### effect of region on rainfall 

```{r}
aov_rainfall <- aov(rainfall~ studycode, data=survey_final_gis)
summary(aov_rainfall)
TukeyHSD(aov_rainfall, conf.level=.95)
```
As p-value for F test of AONVA is very small, rainfall in three regions is not the same. It can be seen the rainfall in region 3 is less than that in the other two.  


### effect of region on wealth 
```{r}
aov_vechval <- aov(vechval~ studycode, data=survey_final_gis)
summary(aov_vechval)
TukeyHSD(aov_vechval, conf.level=.95)

```
The family wealth is significant different in three regions. The wealth in region 3 is the largest.  

# Modelling

## Multiple Logistic Regression Model
First to create a few different MLRs to see what the relationship is between
adaptation methods and water availability.

### Model 1
In this model, I will select the largest amount of variables. The dependent variable will be the general adaptation variable. The independent variables will be:
  - Drainage area
  - Lot size (can't find in data - ask Mariana)
  - Soil type (100% missing - so maybe not)
  - Vehicle value
  - SPI - using the maximum in the year
  - Risk
  - Lot value

(This model did not converge)
  
  
### Model 2
Trying with only drainage area to see if there's a relationship since previous model did not converge
```{r}
fit_2 <- glm(any_adaptation ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_2)
```
The drainage area is not a statistically significant parameter in this case. However, there are outliers in the data that might be affecting this. Let's investigate

```{r}
plot(fit_2, which =4)
```
Fitting the model without observation 206
```{r}
survey_final_gis_no <- survey_final_gis[-c(206),]
fit_3 <- glm(any_adaptation ~ drainage_area_km, family = binomial,data = survey_final_gis_no)
summary(fit_3)
```
```{r}
plot(fit_3, which = 4)
```
Now to check Log Reg assumptions
```{r}
probabilities <- predict(fit_3, type = "response")
# Select only numeric predictors
mydata <- survey_final_gis_no %>%
  dplyr::select_if(is.numeric) %>%
  select(drainage_area_km)
predictors <- colnames(mydata)
# Bind the logit and tidying the data for plot
mydata <- mydata %>%
  mutate(logit = log(probabilities/(1-probabilities))) %>%
  gather(key = "predictors", value = "predictor.value", -logit)
```

```{r}
ggplot(mydata, aes(logit, predictor.value))+
  geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth(method = "loess") + 
  theme_bw() + 
  facet_wrap(~predictors, scales = "free_y")
```


### Model 3 - Transforming Drainage Area
Let's look at the distribution of drainage area to see if we should transform it

```{r}
g <- ggplot(data = survey_final_gis_no, aes(x = log(drainage_area_km)))
g + geom_histogram() + labs(title = "Histogram of log(Drainage Area)")
```

Now to create the model:

```{r}
fit_transform <- glm(any_adaptation ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_transform)
<<<<<<< HEAD

=======
```

### Model 4
For this model, we will take Dr. Harris's suggestion of an indicator variable for
the large drainage area lots (I believe is what he meant)

First going to see the fit with outliers chopped off

```{r}
survey_final_gis_filtered <- survey_final_gis %>% filter(drainage_area_km < 500)
fit_4 <- glm(any_adaptation ~ drainage_area_km, data = survey_final_gis_filtered, family = binomial)
summary(fit_4)
```


### Indivdual Adaptation Variables
>>>>>>> 325ff5089c83501369aed9c3b40bfd7b1b5bfa1f
Now look at individual adaptation variables
```{r}
fit_cattle <- glm(cattle_management ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_cattle)

fit_cattle_log = glm(cattle_management ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_cattle_log)
```
Saw significance when using log(drainage area)

Next lets look at pasture management
```{r}
fit_pasture <- glm(pasture_management ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_pasture)

fit_pasture_log <- glm(pasture_management ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_pasture_log)
```
Similar results, but with a super low p for log(drainage area)

Next lets look at forest conservation
```{r}
fit_forest <- glm(forest_conservation ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_forest)

fit_forest_log <- glm(forest_conservation ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_forest_log)
```

Log(drainage area) is highly significant again

Let's now look at water management
```{r}
fit_water <- glm(water_management ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_forest)

fit_water_log <- glm(water_management ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_forest_log)
```

Similar results again. Log(drainage area) is significant for each of the adaptation categories.

What happens if I filter by "have cattle" and look at cattle_management

```{r}
survey_havecattle <-survey_final_gis %>% filter(havecattle ==1)
fit_cattle_have <- glm(cattle_management ~ drainage_area_km, family = binomial,data = survey_havecattle)
summary(fit_cattle_have)

fit_cattle_have_log <- glm(cattle_management ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_cattle_have_log)
```

Interesting, if we only include those that have cattle, the significance of the log(drainage area) disappears.

Let's see if anything similar happens when filtered by havecattle for any of the other adaptations
```{r}
fit_pasture_have_log <- glm(pasture_management ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_pasture_have_log)

fit_forest_have_log <- glm(forest_conservation ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_forest_have_log)

fit_water_have_log <- glm(water_management ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_forest_have_log)
```

We still have a significant log(drainage area) for all of the other adaptation categories.


Let's also look at deforestation as a linear model
```{r}
fit_deforest <- lm(cleared_area ~ drainage_area_km, data = survey_final_gis)
summary(fit_deforest)

fit_deforest_log <- lm(cleared_area ~ log(drainage_area_km), data = survey_final_gis)
summary(fit_deforest_log)
```

We see significance for both the drainage area and the log(drainage area). The cleared-area is just a total cleared area and not an amount cleared that year, so I don't know what we would actually interpret from this model. Probably not something to include.

Let's repeat some things with SPI and drainage area

```{r}
fit_with_SPI <- glm(any_adaptation ~ log(drainage_area_km), family = binomial, data = survey_final_gis)
summary(fit_with_SPI)

head(survey_final_gis)
```
### modeling for the wealth
```{r}
# remove data of vechval with "NA"
survey_final_gis_wealth <- survey_final_gis %>% filter(!is.na(survey_final_gis$vechval))
fit_wealth_water <- glm(water_management~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_cattle  <- glm(cattle_management~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_pasture  <- glm(pasture_management~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_forest <- glm(forest_conservation~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_any_adaptation <- glm(any_adaptation~  vechval , family = binomial, data = survey_final_gis_wealth )
summary(fit_wealth_water)
summary(fit_wealth_cattle)
summary(fit_wealth_pasture)
summary(fit_wealth_forest)
summary(fit_wealth_any_adaptation )

```
```{r}
fit_lot_water<- glm(water_management ~ lotprice, family = binomial, data = survey_final_gis )
summary(fit_lot_water  )
```

=======
---
title: "st542_project10"
author: "Jingjing Li"
date: "2023-06-15"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(haven)
library(psych)
```
### Exploring the merged data

```{r}
survey_merged <- read_dta("../data/merged_survey_panel96_19_noID.dta") 
dim(survey_merged)
survey_merged_clean <- survey_merged %>% 
  filter(!is.na(lotid_qualtrics))
# Removing duplicates for modelling - forgot to ask Mariana which should be kept, so just keeping first for now
# Also open to just dropping all duplicated lotids - there's only 42
survey_final <- survey_merged_clean %>% distinct(lotid_qualtrics, .keep_all = TRUE)
```

Now to check for nulls in columns:
```{r}
missing_vals <- survey_final %>% summarise_all(~sum(is.na(.)))
t(missing_vals)
```
There seem to be a lot of columns missing 1026 values, which puts the numbers back to what we saw in the original data. Clearly there is a lot of missing data across all areas, so I don't think we should drop all rows missing data. 

## Feature Engineering

### Creating One Drainage Area Column 
I will create one drainage area column, and then will drop all rows that don't have drainage area. Not sure if this is the best approach but this data is so messy!

```{r}
# There are two lots that have areas in each region - will drop them
duplicate_regions <- survey_final %>% filter(!is.na(OPO_area_km) & !is.na(RO_area_km)) %>% select(lotid_qualtrics)
survey_final <- survey_final %>% filter(!(lotid_qualtrics %in% duplicate_regions$lotid_qualtrics))
survey_final <- survey_final %>% mutate(drainage_area_km = ifelse(!is.na(Drainage_AreaKM_Ari), Drainage_AreaKM_Ari, ifelse(!is.na(OPO_area_km), OPO_area_km, ifelse(!is.na(RO_area_km), RO_area_km, NA))),
                                        region = ifelse(studycode == 1, "Ariquemes",
                                                        ifelse(studycode == 2, "Ouro Preto do Oeste", ifelse(studycode == 3, "Rolim de Moura", NA)))) %>%
  select(-c(Drainage_AreaKM_Ari, RO_area_km, OPO_area_km))
```


### Creating Individual Adaptation Method Booleans
First, I need to create the individual adaptation method booleans i.e. `cattle_management` if the farmer employed any cattle management strategy. Then I can combine these to make a general adaptation method boolean.

```{r}
survey_final <- survey_final %>% mutate(cattle_management = case_when(feed_cattle == 1  | soil_analysis == 1 | cattleinputs_breed == 1 | semiconfine == 1 | sellcattle_droughtexp == 1 ~ 1, .default = 0),
                                        pasture_management = case_when(irrigation_pas == 1 | pasture_productivity == 1 | fert_pasture ==1 | pest_pasture == 1 | fallow == 1 ~ 1, .default = 0),
                                        forest_conservation = case_when(forest_rec == 1 | keep_veg == 1 ~ 1, .default = 0),
                                        water_management = case_when(waterstructure == 1 | well_have == 1 | dam_have == 1 | reservoir_have == 1 ~ 1, .default = 0)) # Chose to use case when to account for NA - need to ask Mariana how these should be treated

# Looking at distributions
g <- ggplot(data = survey_final, aes(x = cattle_management))
g + geom_bar() + labs(title = "Bar Plot of Cattle Management")

g <- ggplot(data = survey_final, aes(x = pasture_management))
g + geom_bar() + labs(title = "Bar Plot of Pasture Management")

g <- ggplot(data = survey_final, aes(x = forest_conservation))
g + geom_bar() + labs(title = "Bar Plot of Forest Conservation")

g <- ggplot(data = survey_final, aes(x = water_management))
g + geom_bar() + labs(title = "Bar Plot of Water Management")

```

### Creating General Adaptation Method Boolean
Now that the indivdual adaptation measures have been created, I will create a general adaptation method boolean

```{r}
survey_final <- survey_final %>% mutate(any_adaptation = ifelse(cattle_management == 1 | pasture_management == 1 | forest_conservation == 1 | water_management == 1,1,0))

# Looking at the distribution
g <- ggplot(data = survey_final, aes(x = any_adaptation))
g + geom_bar() + labs(title = "Bar Plot of Any Adaptation")
```
Not a fully even data set, but better balanced than I predicted it would be. I don't
think rebalancing is needed for the final analysis.

Now let's see if we remove farmers who don't have GIS information
```{r}
survey_final_gis <- survey_final %>% filter(!is.na(drainage_area_km))

# Looking at the distribution
g <- ggplot(data = survey_final_gis, aes(x = any_adaptation))
g + geom_bar() + labs(title = "Bar Plot of Any Adaptation")
```
Now it is very imbalanced. Not as horrible as it could be - but could affect the outputs of the model. Is SPI still missing for these?

```{r}
missing_vals_gis <- survey_final_gis %>% summarise_all(~sum(is.na(.)))
t(missing_vals_gis)
```
Now only for 140 of them, who are also missing region. Let's drop those and see:
```{r}
survey_final_gis_clean <- survey_final_gis %>% filter(!is.na(SPImin_year))

# Looking at the distribution
g <- ggplot(data = survey_final_gis_clean, aes(x = any_adaptation))
g + geom_bar() + labs(title = "Bar Plot of Any Adaptation")

# Looking at the distribution for the individual variables
g <- ggplot(data = survey_final_gis_clean, aes(x = cattle_management))
g + geom_bar() + labs(title = "Bar Plot of Cattle Management")

g <- ggplot(data = survey_final_gis_clean, aes(x = pasture_management))
g + geom_bar() + labs(title = "Bar Plot of Pasture Management")

g <- ggplot(data = survey_final_gis_clean, aes(x = forest_conservation))
g + geom_bar() + labs(title = "Bar Plot of Forest Conservation")

g <- ggplot(data = survey_final_gis_clean, aes(x = water_management))
g + geom_bar() + labs(title = "Bar Plot of Water Management")
```
And everyone did some sort of adaptation method. Great - this means if we drop all missing water data, we can't look at a relationship between general adaptation and results. If we include rows that are missing SPI and just use drainage area, there is still a model that can be run.

## Exploratory Data Analysis

Start EDA with some individual variable histograms.

```{r}
g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km))
g + geom_histogram() + labs(title = "Histogram of Drainage Area")

g <- ggplot(data = survey_final_gis, aes(x = rainfall))
g + geom_histogram() + labs(title = "Boxplot of Rainfall")

g <- ggplot(data = survey_final_gis, aes(x = SPI_year))
g + geom_histogram() + labs(title = "Histogram of Average Yearly SPI")

g <- ggplot(data = survey_final_gis, aes(x = vechval))
g + geom_histogram() + labs(title = "Histogram of Vehicle Value")

g <- ggplot(data = survey_final_gis, aes(x = family))
g + geom_histogram() + labs(title = "Histogram of Family Size")

g <- ggplot(data = survey_final_gis, aes(x = lotprice))
g + geom_histogram() + labs(title = "Histogram of Lot Price")

```
Next, we can look at some box plots
```{r}
g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km))
g + geom_boxplot() + labs(title = "Boxplot of Drainage Area")

g <- ggplot(data = survey_final_gis, aes(x = rainfall))
g + geom_boxplot() + labs(title = "Boxplot of Rainfall")

g <- ggplot(data = survey_final_gis, aes(x = SPI_year))
g + geom_boxplot() + labs(title = "Boxplot of Average Yearly SPI")

g <- ggplot(data = survey_final_gis, aes(x = SPI_dry))
g + geom_boxplot() + labs(title = "Boxplot of Average in Peak Dry Season")

g <- ggplot(data = survey_final_gis, aes(x = SPI_wet))
g + geom_boxplot() + labs(title = "Boxplot of Average in Peak Wet Season")

g <- ggplot(data = survey_final_gis, aes(x = vechval))
g + geom_boxplot() + labs(title = "Boxplot of Vehicle Value")

g <- ggplot(data = survey_final_gis, aes(x = family))
g + geom_boxplot() + labs(title = "Boxplot of Family Size")

g <- ggplot(data = survey_final_gis, aes(x = lotprice))
g + geom_boxplot() + labs(title = "Boxplot of Lot Price")


```


Let's look at boxplots of drainage area and vehicle value (wealth) grouped by cattle.

```{r}
g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km, y = as.factor(havecattle)))
g + geom_boxplot() + labs(title = "Boxplot of Drainage Area grouped by Cattle") +
  ylab("Cattle")

g <- ggplot(data = survey_final_gis, aes(x = vechval, y = as.factor(havecattle)))
g + geom_boxplot() + labs(title = "Boxplot of Vehicle Value grouped by Cattle") +
  ylab("Cattle")

```

Let's look at a few box plots split by region

```{r}

g <- ggplot(data = survey_final_gis, aes(x = drainage_area_km, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Drainage Area grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = SPI_dry, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Dry Season SPI grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = SPI_wet, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Wet Season SPI grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = rainfall_dry, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Dry Season Rainfall grouped by Region") +
  ylab("Region")

g <- ggplot(data = survey_final_gis, aes(x = rainfall_wet, y = as.factor(studycode)))
g + geom_boxplot() + labs(title = "Boxplot of Wet Season Rainfall grouped by Region") +
  ylab("Region")

```

Next lets look at pair plots

```{r}
pairs_subset <- survey_final_gis %>% select(c(drainage_area_km, SPImax_year, rainfall, studycode, havecattle, vechval, lotprice, incbeef))
pairs(pairs_subset)

```

Let's look at how many households fall into each region

```{r}
table(survey_final_gis$studycode)
```

Let's also look at some basic summary statistics

```{r}
describe(survey_final_gis[c('drainage_area_km', 'rainfall', 'rainfall_wet', 'rainfall_dry', 'SPI_wet', 'SPI_dry', 'incbeef', 'vechval', 'lotprice', 'family')], fast=TRUE)

```


I'm curious about how many farms made multiple adaptations

```{r}
table(survey_final_gis$cattle_management, survey_final_gis$pasture_management)

table(survey_final_gis$cattle_management, survey_final_gis$forest_conservation)

table(survey_final_gis$cattle_management, survey_final_gis$water_management)

table(survey_final_gis$pasture_management, survey_final_gis$forest_conservation)

table(survey_final_gis$pasture_management, survey_final_gis$water_management)

table(survey_final_gis$forest_conservation, survey_final_gis$water_management)
```

It looks like the majority of farmers who adapted did more than one thing.

# Statisitcal analysis

## t test for milkincome
We would like to see if there is any difference of milkincome between dry and wet season. The null hypothesis is mean value of milkincome is the same for dry and wet season. H0: $\mu_{mlkinc,dry}=\mu_{mlkinc,wet)}$ vs Ha: $\mu_{mlkinc,dry} \ne \mu_{mlkinc,wet)}$  

```{r}
t.test(survey_final_gis_clean$milkincwet,survey_final_gis$milkincdry,na.rm=TRUE,var.equal = FALSE)
```
As p is much less than 0.05, we are confident the mean value of milkincome of wetseason is significant higher than dry season.  

# t-test for SPI 
We would like to see if there is any difference in SPI between dry and wet season. The null hypothesis is mean value of SPI is the same for dry and wet season. H0: $\mu_{SPI,dry}=\mu_{SPI,wet)}$ vs Ha: $\mu_{SPI,dry} \ne \mu_{SPI,wet)}$ 
```{r}
t.test(survey_final_gis_clean$SPI_wet6,survey_final_gis$SPI_dry6,na.rm=TRUE,var.equal = FALSE)
t.test(survey_final_gis_clean$SPI_wet,survey_final_gis$SPI_dry,na.rm=TRUE,var.equal = FALSE)
```
Average SPI and average monthly SPI in the peak of wet  season is significantly higher than those of dry season.   

## Anova analysis
We hypothesize drainage area of three regions may be different. We propose null hypothesis is drainage area means of three regions are the same. The alternativae hypothesis is drainage area means of three regions are different.  

### effect of region on drainage_area_km
```{r}
survey_final_gis$studycode<- factor(survey_final_gis$studycode)
aov_drainage <- aov(drainage_area_km ~ studycode, data=survey_final_gis)

summary(aov_drainage)
TukeyHSD(aov_drainage, conf.level=.95)
```
As p-value of F test is close 0.05,  the drainage area means of three regions are somehow different. We further compared drainage means of three regions and found region3 (Rolim) has the highest drainage area, OPO has the smallest.  

### effect of region on income from beef
We then wants to see if there is difference in income from beef at different regions.
```{r}
aov_incbeef <- aov(incbeef ~ studycode, data=survey_final_gis)
summary(aov_incbeef)
TukeyHSD(aov_incbeef, conf.level=.95)
```
As p value is larger than 0.05, we can not reject our hypothesis that the means of income from beef at different regions are the same.  


### effect of region on rainfall 

```{r}
aov_rainfall <- aov(rainfall~ studycode, data=survey_final_gis)
summary(aov_rainfall)
TukeyHSD(aov_rainfall, conf.level=.95)
```
As p-value for F test of AONVA is very small, rainfall in three regions is not the same. It can be seen the rainfall in region 3 is less than that in the other two.  


### effect of region on wealth 
```{r}
aov_vechval <- aov(vechval~ studycode, data=survey_final_gis)
summary(aov_vechval)
TukeyHSD(aov_vechval, conf.level=.95)

```
The family wealth is significant different in three regions. The wealth in region 3 is the largest.  

# Modelling

## Multiple Logistic Regression Model
First to create a few different MLRs to see what the relationship is between
adaptation methods and water availability.

### Model 1
In this model, I will select the largest amount of variables. The dependent variable will be the general adaptation variable. The independent variables will be:
  - Drainage area
  - Lot size (can't find in data - ask Mariana)
  - Soil type (100% missing - so maybe not)
  - Vehicle value
  - SPI - using the maximum in the year
  - Risk
  - Lot value

(This model did not converge)
  
  
### Model 2
Trying with only drainage area to see if there's a relationship since previous model did not converge
```{r}
fit_2 <- glm(any_adaptation ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_2)
```
The drainage area is not a statistically significant parameter in this case. However, there are outliers in the data that might be affecting this. Let's investigate

```{r}
plot(fit_1, which = 1)
```

Now look at individual adaptation variables
```{r}
fit_cattle <- glm(cattle_management ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_cattle)

fit_cattle_log = glm(cattle_management ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_cattle_log)
```
Saw significance when using log(drainage area)

Next lets look at pasture management
```{r}
fit_pasture <- glm(pasture_management ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_pasture)

fit_pasture_log <- glm(pasture_management ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_pasture_log)
```
Similar results, but with a super low p for log(drainage area)

Next lets look at forest conservation
```{r}
fit_forest <- glm(forest_conservation ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_forest)

fit_forest_log <- glm(forest_conservation ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_forest_log)
```

Log(drainage area) is highly significant again

Let's now look at water management
```{r}
fit_water <- glm(water_management ~ drainage_area_km, family = binomial,data = survey_final_gis)
summary(fit_forest)

fit_water_log <- glm(water_management ~ log(drainage_area_km), family = binomial,data = survey_final_gis)
summary(fit_forest_log)
```

Similar results again. Log(drainage area) is significant for each of the adaptation categories.

What happens if I filter by "have cattle" and look at cattle_management

```{r}
survey_havecattle <-survey_final_gis %>% filter(havecattle ==1)
fit_cattle_have <- glm(cattle_management ~ drainage_area_km, family = binomial,data = survey_havecattle)
summary(fit_cattle_have)

fit_cattle_have_log <- glm(cattle_management ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_cattle_have_log)
```

Interesting, if we only include those that have cattle, the significance of the log(drainage area) disappears.

Let's see if anything similar happens when filtered by havecattle for any of the other adaptations
```{r}
fit_pasture_have_log <- glm(pasture_management ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_pasture_have_log)

fit_forest_have_log <- glm(forest_conservation ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_forest_have_log)

fit_water_have_log <- glm(water_management ~ log(drainage_area_km), family = binomial,data = survey_havecattle)
summary(fit_forest_have_log)
```

We still have a significant log(drainage area) for all of the other adaptation categories.


Let's also look at deforestation as a linear model
```{r}
fit_deforest <- lm(cleared_area ~ drainage_area_km, data = survey_final_gis)
summary(fit_deforest)

fit_deforest_log <- lm(cleared_area ~ log(drainage_area_km), data = survey_final_gis)
summary(fit_deforest_log)
```

We see significance for both the drainage area and the log(drainage area). The cleared-area is just a total cleared area and not an amount cleared that year, so I don't know what we would actually interpret from this model. Probably not something to include.

Let's repeat some things with SPI and drainage area

```{r}
fit_with_SPI <- glm(any_adaptation ~ log(drainage_area_km), family = binomial, data = survey_final_gis)
summary(fit_with_SPI)

head(survey_final_gis)
```
### modeling for the wealth
```{r}
# remove data of vechval with "NA"
survey_final_gis_wealth <- survey_final_gis %>% filter(!is.na(survey_final_gis$vechval))
fit_wealth_water <- glm(water_management~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_cattle  <- glm(cattle_management~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_pasture  <- glm(pasture_management~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_forest <- glm(forest_conservation~  vechval , family = binomial, data = survey_final_gis_wealth )
fit_wealth_any_adaptation <- glm(any_adaptation~  vechval , family = binomial, data = survey_final_gis_wealth )
summary(fit_wealth_water)
summary(fit_wealth_cattle)
summary(fit_wealth_pasture)
summary(fit_wealth_forest)
summary(fit_wealth_any_adaptation )

```
```{r}
fit_lot_water<- glm(water_management ~ lotprice, family = binomial, data = survey_final_gis )
summary(fit_lot_water  )
```

>>>>>>> 5943cb226095ba196c7ef7c2a597565fdd326e4d

# Tree modeling
```{r}
library(tidyverse)
library(caret)
library(rpart)
#split into training and test
train <- sample(1:nrow(survey_final_gis_wealth), size = nrow(survey_final_gis_wealth)*0.8)
test <- dplyr::setdiff(1:nrow(survey_final_gis_wealth), train)
survey_final_gis_Train <- survey_final_gis_wealth[train, ]
survey_final_gis_Test <- survey_final_gis_wealth[test, ]

#fit models using repeated CV


classTree <- train(cattle_management ~  log(drainage_area_km) + vechval , data = survey_final_gis_Train, method = "rpart",
trControl = trainControl(method = "repeatedcv", number = 10, repeats = 5),
preProcess = c("center", "scale"))
baggedTree <- train(cattle_management ~ log(drainage_area_km) + vechval, data = survey_final_gis_Train, method = "treebag",
trControl = trainControl(method = "repeatedcv", number = 10, repeats = 5),
preProcess = c("center", "scale"))
rfTree <- train(cattle_management ~ log(drainage_area_km) + vechval, data = survey_final_gis_Train, method = "rf",
trControl = trainControl(method = "repeatedcv", number = 10, repeats = 5),
preProcess = c("center", "scale"))
boostTree <- train(cattle_management ~ log(drainage_area_km) + vechval, data = survey_final_gis_Train, method = "gbm",
trControl = trainControl(method = "repeatedcv", number = 10, repeats = 5),
preProcess = c("center", "scale"), verbose = FALSE)

treeTbl <- table(data.frame(pred = predict(classTree, survey_final_gis_Test ), true = survey_final_gis_Test $cattle_management))
baggedTbl <- table(data.frame(pred = predict(baggedTree, survey_final_gis_Test ), true = survey_final_gis_Test $cattle_management))
rfTbl <- table(data.frame(pred = predict(rfTree, survey_final_gis_Test ), true = survey_final_gis_Test $cattle_management))
boostTbl <- table(data.frame(pred = predict(boostTree, survey_final_gis_Test), true = survey_final_gis_Test $cattle_management))
#misclassificatoon rate
1-c(tree = sum(diag(treeTbl)/sum(treeTbl)), bag = sum(diag(baggedTbl)/sum(baggedTbl)),
rf = sum(diag(rfTbl)/sum(rfTbl)),
boost = sum(diag(boostTbl)/sum(boostTbl)))
```

